Very Preliminary Draft

### 2021-07-05 article/draft.tex
upadate

\documentclass{article}
%\documentclass[a4paper,12pt,twoside]{book}
\usepackage{amsmath}
\usepackage{cite}
%\usepackage{tikz}
\usepackage{bm}
\usepackage{tikz,tkz-tab}
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{hyperref}
\usepackage{mathtools}
%\usepackage{subcaption}
\usepackage{color}
\usepackage{setspace}
\usepackage{empheq}
\usepackage{bbm, dsfont}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{geometry}
\usepackage{enumitem} 
\usepackage[bottom]{footmisc}
\usetikzlibrary{arrows}
\usepackage{lscape}
\usepackage{tcolorbox}
\usepackage{caption}
%\usepackage{graphicx}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{cite}
\usetikzlibrary{shapes,snakes}
\renewcommand{\labelitemi}{$\bullet$}
\renewcommand{\labelitemii}{$\diamond$}

@@

\begin{document}
\title{Very Preliminary Draft}
\maketitle

@@

In this article, we focus on MSMP's policies and actions regarding content with low credibility or false information, commonly referred to as {\it Fake News}.\footnote{For an overview on the {\it concept} of {\it Fake News}, we refer the reader to the article The science of fake news by Lazer et al. (2018)~\cite{lazer}.} The {\it Fake News} phenomenon is still ill-defined by the academic community as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience. The rise of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. {\color{red} Fact-checkers + no laws or not so much.}

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content.\footnote{For Facebook, Twitter and Youtube see respectively the following updates: } As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of MSMP to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. {\color{pink} Say what they say they do + partnerships with fact-checkers + algorithms + community, Make point about ``recycling" existing policies (e.g. to tackle terrorism) and apply it to misinformation+ idea about academic community needs to be able to study this phenomenon and assess the impact of the policies, what effects they have etc.}

@@

\bigskip

%{\color{red} Policies not specific to misinformation, but to enforce laws + things already used for terrorism and hate speech. Intro or/and Discussion. accounts: hard to say suspension for which.}

%Content can be withheld in compliance to specific laws of each country, ex: ..;
% Misinformation : which law ? not defined !

@@

 %We We are aware that other APIs or databases can be useful on that matter, but we will only mention here the ones about which we have some experience.

%CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages. BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms.

%Section 230 in the US + EU 
%Pressure to moderate and mix-up of platforms with public square… 
%Rôle dans la production/médiation : P-M-R 
%Pourquoi faut-il vérifier leurs politiques ? D’une part parce que ce sont les plateformes elles-mêmes qui annoncent leurs effets. D’autre part beaucoup de choses indirectes. 
%
%Add disclaimer in a footnote early on in the document : document not meant to be an exhaustive overview of all the existing regulations and point to three or four main legal sources for exhaustive references.

%Such policies include the introduction of notices (Twitter, see link) or flags (Facebook, see…) to signal to users inaccurate or false content, the temporary or permanent suspension of accounts (Twitter) or pages (Facebook), reducing the visibility of posts or pages (Facebook), or blocking users from sharing specific URL links (see link). The objective of this article is twofold. First, we study the impact of a number of  implemented measures by mainstream platforms to tackle (mis)disinformation.

@@

Main stream social media platforms such as Facebook, Twitter and Youtube, may suspend the account of a specific user when they deem that the platforms' rules have been violated. Account suspension can be temporary or permanent.\footnote{A list of notable Twitter temporary and permanent suspensions can be found on wikipedia: \href{https://en.wikipedia.org/wiki/Twitter_suspensions}{https://en.wikipedia.org/wiki/Twitter\_suspensions}.}  When the suspension is temporary the user is prohibited for a limited period of time from posting content on their account, but created content prior to suspension remains available to the user and their followers. However, when the suspension is permanent, in most cases, followers or subscribers have no longer access to the content prior to the suspension and the user can no longer use the account to create new content.  In what follows, we focus on the implementation of this policy by several platforms and provide simple examples to illustrate.

%\begin{figure}[h]
%\centering
%\includegraphics[scale=0.2]{./img/fig1.png}  \hfill \includegraphics[scale=0.2]{./img/crg.png} \hfill \includegraphics[scale=0.2]{./img/life.png} 
%\caption{Screenshots taken on June 14, 2021 at 4pm UTC. }
%\label{fig1}
%\end{figure}

@@

When an account is permanently suspended by Facebook, it disappears from the platform (so its data cannot be scrapped anymore), and its data also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.}

Interestingly, Facebook is regularly publishing a monthly {\it coordinated inauthentic behavior} report where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged.\footnote{See the April 2021 report for an example: \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{ (see the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used the minet command line tool \cite{minet} to collect data. We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

\begin{figure}
	\centering
	\includegraphics[scale=0.35]{./img/fb/fig1_fb.png}
	\caption{Number of Facebook posts published each day by the Facebook page {\it Donald J. Trump} between January $1$, $2020$ and June $15$, $2021$. The data corresponds to $6$ $083$ posts retrieved from the CrowdTangle API using the {\it posts} endpoint.}
	\label{fig1_fb}
\end{figure}

@@

{\color{red} N.B. Manual decision, not the automatic strike system}
\begin{figure}
\centering
\includegraphics[scale=0.35]{./img/lifesite.jpg} 
\includegraphics[scale=0.35]{./img/lifesite_rolling_7_lifesite.jpg}
\caption{Panel (a): number of Tweets per day of the Twitter account $@Lifesite$ linked to the website lifesitenews.com from January, 2019 until April 2021. Panel (b): number of Tweets per day that have shared a lifesitenews.com URL link from January, 2019 until April 2021. }
\label{fig2}
\end{figure}

In this section, we discuss one case of temporary suspension on Twitter.\footnote{See the official documentation on the Twitter's Help Center regarding account suspension: \href{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts}{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts }.} The Twitter account of the website lifesitenews.com has been suspended for at least two periods of time: from end of 2019 until fall 2020 for 308 days, then again since January 2021 for having violated Twitter Rules\footnote{See Lifesitenews's article discussing the reason for the suspension: \href{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}. Twitter rules can be found at: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{https://help.twitter.com/en/rules-and-policies/twitter-rules}. }. In particular, this website has several failed fact-checks concerning the published articles, according to Iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/life-site-news/}{https://mediabiasfactcheck.com/life-site-news/}}. We collected the activity (tweets, replies, quotes, retweets) on their Twitter account $@LifeSite$ via the Twitter API, using the historical search endpoint.\footnote{See the documentation: \href{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}.} We then plotted the number of Tweets, Retweets, Quotes and Replies per day, as shown in panel $a$ of figure \ref{fig2}). The two periods of temporary suspension are clearly observed in the data as the user(s) of the account were not allowed to use the functionalities of the Twitter Platform.

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.25]{./img/oann/fig1_oann.png}
	\includegraphics[scale=0.25]{./img/oann/fig2_oann.png} 
	\caption{panel (a): Number of Youtube videos uploaded each day by the youtube channel {\it One America news Network} November 1, 2020 and January 1, 2021. Panel (b): accumulated view counts for videos. The metrics correspond to the videos’  publishing date and the data is retrieved from the youtube API with the {\it playlists} and  {\it videos} endpoints. }
	\label{fig1_oann}
\end{figure}

First, we investigate the temporary suspension case of the Youtube Channel of {\it One America News channel}. This channel received a first strike on November 24, 2020 for the promotion of  a false cure for COVID19.\footnote{Reference?} We collected the activity of the channel OANN (video counts, view counts) using  the Youtube API v3, between November 2020 and January 2021. For the video counts, we used the playlist endpoint to retrieve the videos uploaded with their publishing date and for the view count we used the IDs of the videos we had from the playlists and via the videos endpoint we retrieved the view counts on June 2021.\footnote{See the Google documentation  \href{https://developers.google.com/youtube/v3/docs/videos/list}{https://developers.google.com/youtube/v3/docs/videos/list} and \href{https://developers.google.com/youtube/v3/docs/playlists/list}{https://developers.google.com/youtube/v3/docs/playlists/list}}

@@

In addition, as shown in figure \ref{fig1_oann} when comparing the week before the suspension from 2020/11/17 to 2020/11/23 and one week after from 2020/12/01 to 2020/12/07 it was found that the view count had decreased by ($-45\%$ or $-60\%$), even though the posted number of videos was similar (before $81$ videos; after 93 videos or 86 videos). In other words, the suspension period may have a good impact on reducing the audience interest or reach to the channel. Besides that, OANN decided to move officially to Rumble on March 17, 2021 as announced on their Twitter account (see figure \ref{fig2_oann}) and their upload activity on their Youtube channel is close to zero since that announcement.

\begin{figure}
	\centering
	\includegraphics[scale=0.2]{./img/oann/fig3_oann.png}
	\includegraphics[scale=0.3]{./img/tony/fig3_tony.png}
	\caption{Panel (a): Tweet announcing moving to rumble by OANN (Twitter), Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Panel (b): Tony Heller's tweet after getting suspended from Youtube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{fig2_oann}
\end{figure}

@@

We now turn to our second example, the temporary suspension of the Youtube channel Tony Heller. This channel got its first strike after posting a video about an anti-covid-lockdown doctor getting arrested (see screenshot in figure \ref{fig2_oann}). The suspension period was for one week from September 29 until October 5. We applied the same methods as in the previous example for the data collection. Figure \ref{fig2_oann} shows the daily number of videos uploaded by the channel. The suspension period can be observed clearly in the historical data of the channel. Moreover, observing the reach of the audience using view counts one week before the suspension starting from 2020/09/22 to 2020/09/28 and one week after the suspension from 2020/10/06 to 2020/10/12 the channel witnessed a drop of $-65\%$ and the videos published in the channel were less by $-16\%$.

\begin{figure}
	\centering
	\includegraphics[scale=0.3]{./img/tony/fig1_tony.png}
	\includegraphics[scale=0.3]{./img/tony/fig2_tony.png}
	\caption{Panel (a): Number of Youtube videos uploaded each day by the Youtube channel {\it Tony Heller} between September 1, 2020 and November 15, 2020. Panel (b): accumulated view counts for videos uploaded by the same Youtube channel. The date corresponds to the videos’  publishing date. 
}
	\label{fig1_tony}
\end{figure}

@@

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through ranking. Facebook ranks each post and ad with a relevancy score, a high score leading to a high likelihood of the post or the ad appearing in someone’s newsfeed. Facebook can thus make a post or a whole account less visible by decreasing the rank of its content [12]. Such {\it reduce} action should be visible in the engagement metrics of the posts: if a post reaches less users because of a lower ranking, it is less likely to receive likes, comments and shares.

Owned by Alex Jones, Infowars is usually described as a misinformation website: it is on the Wikimedia global spam blacklist [8], was classified among websites that have posted deceptive content according to FactCheck.org [9], and its factual reporting has been rated {\it very low} by the Media Bias / Fact Check resource [10]. On May 2, 2019, Facebook announced they would prohibit users from sharing Infowars content unless they are explicitly condemning the material [7].

To verify the measure, we used the CrowdTangle API at the ‘/posts/search’ endpoint (see the documentation [11] for more details) to collect the 37,242 Facebook public posts that shared a link containing ‘infowars.com’ and that were published between January 1, 2019 and December 31, 2020.\footnote{We found in the collected data some Facebook posts that did not directly share an Infowars link (but rather a YouTube or Facebook video containing an Infowars link in its description), thus we excluded such posts from our data to keep only the 27,721 posts directly sharing an Infowars link.} The command used can be found \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_infowars_data.sh}{here}.

The number of public posts sharing an Infowars link remained globally stable in 2019 (Figure X4 top panel), thus Facebook’s announced ‘ban’ did not appear to prevent users from sharing an Infowars link. Interestingly a clear drop in engagement was observed on May 2, 2019 (Figure X4 bottom panel). The number of reactions per post decreased by -94\% when we compared the two months before and after May 2, 2019, the shares decreased by -96\%, and the comments by -93\%.

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.3]{./img/globalresearch_2021-06-14.png}
	\includegraphics[scale=0.3]{./img/globalresearch_14_06_2021_16pm_UTC.png} 
	\caption{Screenshots taken on June 14, 2021. }
	\label{fig3}
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{./img/globalresearch/sum_globalresearch.ca_6_months.jpg} 
\includegraphics[scale=0.3]{./img/globalresearch/engagement_sum_rolling_1_globalresearch.ca.jpg}
\includegraphics[scale=0.3]{./img/globalresearch/tweet.png} 
\includegraphics[scale=0.3]{./img/globalresearch/warning.png}
%\includegraphics[scale=0.3]{./img/globalresearch/sum_rolling_7_globalresearch.jpg} 
%\includegraphics[scale=0.3]{./img/globalresearch/sum_rolling_7_globalresearchall_time.jpg}
\caption{}
\label{fig4}
\end{figure}

@@

{\color{pink} What about something about recommendations ? authoritative content  }
\subsection{Flags and Notices}

@@

>>>>

\documentclass{article}
%\documentclass[a4paper,12pt,twoside]{book}
\usepackage{amsmath}
\usepackage{cite}
%\usepackage{tikz}
\usepackage{bm}
\usepackage{tikz,tkz-tab}
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{hyperref}
\usepackage{mathtools}
%\usepackage{subcaption}
\usepackage{color}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{empheq}
\usepackage{bbm, dsfont}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{geometry}
\usepackage{enumitem} 
\usepackage[bottom]{footmisc}
\usetikzlibrary{arrows}
\usepackage{lscape}
\usepackage{tcolorbox}
\usepackage{caption}
%\usepackage{graphicx}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{cite}
\usetikzlibrary{shapes,snakes}
\renewcommand{\labelitemi}{$\bullet$}
\renewcommand{\labelitemii}{$\diamond$}

@@

\begin{document}

\title{Very Preliminary Draft}

\maketitle

@@

In this article, we focus on MSMP's policies and actions regarding content with low credibility or false information, commonly referred to as {\it Fake News}.\footnote{For an overview on the {\it concept} of {\it Fake News}, we refer the reader to the article The science of fake news by Lazer et al. (2018)~\cite{lazer}.} The {\it Fake News} phenomenon is still ill-defined by the academic community as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience. The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms.

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content.\footnote{For Facebook, Twitter and Youtube see respectively the following updates: } As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of MSMP to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. {\color{pink} Say what they say they do + partnerships with fact-checkers + algorithms + community,  idea about academic community needs to be able to study this phenomenon and assess the impact of the policies, what effects they have etc.}

@@

%\bigskip

%{\color{red} Policies not specific to misinformation, but to enforce laws + things already used for terrorism and hate speech. Intro or/and Discussion. accounts: hard to say suspension for which.} 
% Misinformation : which law ? not defined !

@@

 %We We are aware that other APIs or databases can be useful on that matter, but we will only mention here the ones about which we have some experience.
%CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages. BuzzSumo is a commercial content database that tracks the volume of user interactions with internet content on Facebook, Twitter, and other social media platforms. 
%Section 230 in the US + EU 
%Pressure to moderate and mix-up of platforms with public square… 
%Rôle dans la production/médiation : P-M-R 
%Pourquoi faut-il vérifier leurs politiques ? D’une part parce que ce sont les plateformes elles-mêmes qui annoncent leurs effets. D’autre part beaucoup de choses indirectes. 
%Add disclaimer in a footnote early on in the document : document not meant to be an exhaustive overview of all the existing regulations and point to three or four main legal sources for exhaustive references. 
%Such policies include the introduction of notices (Twitter, see link) or flags (Facebook, see…) to signal to users inaccurate or false content, the temporary or permanent suspension of accounts (Twitter) or pages (Facebook), reducing the visibility of posts or pages (Facebook), or blocking users from sharing specific URL links (see link). The objective of this article is twofold. First, we study the impact of a number of  implemented measures by mainstream platforms to tackle (mis)disinformation.

@@

Mainstream social media platforms may suspend the account of a specific user when they deem that the platforms' rules have been violated. Account suspension can be temporary or permanent.\footnote{A list of notable Twitter temporary and permanent suspensions can be found on wikipedia: \href{https://en.wikipedia.org/wiki/Twitter_suspensions}{https://en.wikipedia.org/wiki/Twitter\_suspensions}.}  When the suspension is temporary the user is prohibited for a limited period of time from posting content on their account, but created content prior to suspension remains available to the user and their followers. However, when the suspension is permanent, in most cases, followers or subscribers have no longer access to the content prior to the suspension and the user can no longer use the account to create new content.  In what follows, we focus on the implementation of this policy by several platforms and provide simple examples to illustrate.

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged.\footnote{See the April 2021 report for an example: \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{ (see the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used the minet command line tool \cite{minet} to collect data.\footnote{The exact command can be found : \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_trump_data.sh}{here}.} We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

\begin{figure}
	\centering
		\begin{multicols}{1}
			\includegraphics[scale=0.2]{./img/fb/fig1_fb.png}
		\end{multicols}
	\caption{Number of Facebook posts published each day by the Facebook page {\it Donald J. Trump} between January $1$, $2020$ and June $15$, $2021$. The data corresponds to $6$ $083$ posts retrieved from the CrowdTangle API using the {\it posts} endpoint.}
	\label{fig1_fb}
\end{figure}

@@

\begin{figure}
\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/lifesite.jpg} 
	\end{multicols}
	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/lifesite_rolling_7_lifesite.jpg}
	\end{multicols}
\caption{Panel (a): number of Tweets per day of the Twitter account $@Lifesite$ linked to the website lifesitenews.com from January, 2019 until April 2021. Panel (b): number of Tweets per day that have shared a lifesitenews.com URL link from January, 2019 until April 2021. }
\label{fig2}
\end{figure}

%paragraph strike system + say following example is a manual decision. 
Twitter has implemented a strike system as part of their Civi Integrity Policy\footnote{See \href{https://help.twitter.com/en/rules-and-policies/election-integrity-policy}{help.twitter.com/en/rules-and-policies/election-integrity-policy}} and their COVID19 misleading information policy.\footnote{See \href{https://help.twitter.com/en/rules-and-policies/medical-misinformation-policy}{help.twitter.com/en/rules-and-policies/medical-misinformation-policy} and \href{https://blog.twitter.com/en\_us/topics/company/2021/updates-to-our-work-on-covid-19-vaccine-misinformation}{blog.twitter.com/en\_us/topics/company/2021/updates-to-our-work-on-covid-19-vaccine-misinformation}} Violations of both policies can entail strikes, where two strikes lead to a 12-hour account lock and five or more strikes lead to permanent suspension from the platform. The 12-hour account lock is hard to observe in the data, especially for users who do not have an over the clock tweeting activity. In this section, we provide one example of a temporary suspension\footnote{See the official documentation on the Twitter's Help Center regarding account suspension: \href{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts}{https://help.twitter.com/en/managing-your-account/suspended-twitter-accounts }.} of a Twitter account, that seems to be the result of a manual decision concerning a Tweet which violated the rules.

The Twitter account of the website lifesitenews.com has been suspended for at least two periods of time: from end of 2019 until fall 2020 for 308 days, then again since January 2021 for having violated Twitter Rules\footnote{See Lifesitenews's article discussing the reason for the suspension: \href{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}{https://www.lifesitenews.com/news/lifesite-is-dumping-twitter-and-so-should-you}. Twitter rules can be found at: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{https://help.twitter.com/en/rules-and-policies/twitter-rules}. }. In particular, this website has several failed fact-checks concerning the published articles, according to Iffy.news.\footnote{See \href{https://mediabiasfactcheck.com/life-site-news/}{https://mediabiasfactcheck.com/life-site-news/}}. We collected the activity (tweets, replies, quotes, retweets) on their Twitter account $@LifeSite$ via the Twitter API, using the historical search endpoint.\footnote{See the documentation: \href{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}{https://developer.twitter.com/en/docs/twitter-api/tweets/search/api-reference/get-tweets-search-all}.} We then plotted the number of Tweets, Retweets, Quotes and Replies per day, as shown in panel $a$ of figure \ref{fig2}). The two periods of temporary suspension are clearly observed in the data as the user(s) of the account were not allowed to use the functionalities of the Twitter Platform.

@@

\begin{figure}
	\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.25]{./img/oann/fig1_oann.png}
	\end{multicols}
	\begin{multicols}{1}
		\includegraphics[scale=0.25]{./img/oann/fig2_oann.png} 
	\end{multicols}
	\caption{panel (a): Number of Youtube videos uploaded each day by the youtube channel {\it One America news Network} November 1, 2020 and January 1, 2021. Panel (b): accumulated view counts for videos. The metrics correspond to the videos’  publishing date and the data is retrieved from the youtube API with the {\it playlists} and  {\it videos} endpoints. }
	\label{fig1_oann}
\end{figure}

First, we investigate the temporary suspension case of the Youtube Channel of {\it One America News channel}. This channel received a first strike on November 24, 2020 for the promotion of  a false cure for COVID19.\footnote{See nbcnews,  ``YouTube suspends OANN for violating its Covid-19 policy''. nbcnews, Ahiza García-Hodges, 24 11 2020.} We collected the activity of the channel OANN (video counts, view counts) using  the Youtube API v3, between November 2020 and January 2021. For the video counts, we used the playlist endpoint to retrieve the videos uploaded with their publishing date and for the view count we used the IDs of the videos we had from the playlists and via the videos endpoint we retrieved the view counts on June 2021.\footnote{See the Google documentation  \href{https://developers.google.com/youtube/v3/docs/videos/list}{https://developers.google.com/youtube/v3/docs/videos/list} and \href{https://developers.google.com/youtube/v3/docs/playlists/list}{https://developers.google.com/youtube/v3/docs/playlists/list}}

@@

In addition, as shown in figure \ref{fig1_oann} when comparing the month before the suspension from 2020/10/24 to 2020/11/24 and one month after from 2020/12/01 to 2021/01/01 it was found that the view count decreased by -73\% and the videos uploaded by -55\%. Besides that, OANN decided to move officially to Rumble on March 17, 2021 as announced on their Twitter account (see figure \ref{fig2_oann}) and their upload activity on their Youtube channel is close to zero since that announcement.

\begin{figure}
	\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/oann/fig3_oann.png}
	\end{multicols}

	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/tony/fig3_tony.png}
	\end{multicols}

	\caption{Panel (a): Tweet announcing moving to rumble by OANN (Twitter), Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Panel (b): Tony Heller's tweet after getting suspended from Youtube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{fig2_oann}
\end{figure}

@@

We now turn to our second example, the temporary suspension of the Youtube channel Tony Heller. This channel got its first strike after posting a video about an anti-covid-lockdown doctor getting arrested (see screenshot in figure \ref{fig2_oann}). The suspension period was for one week from September 29 until October 5. We applied the same methods as in the previous example for the data collection. Figure \ref{fig2_oann} shows the daily number of videos uploaded by the channel. The suspension period can be observed clearly in the historical data of the channel.  observing the reach of the audience 
Figure 3: Tony heller tweet after getting suspended from youtube (Twitter)
using view counts one month before the suspension starting from 2020/08/28 to 2020/09/28 and one month after the suspension from 2020/10/05 to 2020/11/05 the channel witnessed a drop of view counts by $-69.5\%$ and the videos published in the channel were less by $-29\%$. This drop in views can show that the  suspension  period
may  have  a  good  impact  on  reducing  the  audience interest or reach to the channel.

\begin{figure}
	\centering
		\begin{multicols}{1}
			\includegraphics[scale=0.3]{./img/tony/fig1_tony.png}
		\end{multicols}

		\begin{multicols}{1}
			\includegraphics[scale=0.3]{./img/tony/fig2_tony.png}
		\end{multicols}
	\caption{Panel (a): Number of Youtube videos uploaded each day by the Youtube channel {\it Tony Heller} between September 1, 2020 and November 15, 2020. Panel (b): accumulated view counts for videos uploaded by the same Youtube channel. The date corresponds to the videos’  publishing date. 
}
	\label{fig1_tony}
\end{figure}

@@

One of Facebook’s measures to regulate misinformation is to reduce the spread of misleading content through ranking. Facebook ranks each post and ad with a relevancy score, a high score leading to a high likelihood of the post or the ad appearing in someone’s newsfeed. Facebook can thus make a post or a whole account less visible by decreasing the rank of its content.\footnote{Lyons, T. (2018, May 22). The three-part recipe for cleaning up your news feed. Facebook Newsroom. \href{https://about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}{about.fb.com/news/2018/05/inside-feed-reduce-remove-inform/}.} Such {\it reduce} action should be visible in the engagement metrics of the posts: if a post reaches less users because of a lower ranking, it is less likely to receive likes, comments and shares.

Owned by Alex Jones, Infowars is usually described as a misinformation website: it is on the Wikimedia global spam blacklist\footnote{See \href{https://meta.wikimedia.org/wiki/Spam\_blacklist}{meta.wikimedia.org/wiki/Spam\_blacklist}.}, was classified among websites that have posted deceptive content according to FactCheck.org\footnote{See \href{https://www.factcheck.org/2017/07/websites-post-fake-satirical-stories/}{https://www.factcheck.org/2017/07/websites-post-fake-satirical-stories/}.}, and its factual reporting has been rated {\it very low} by the Media Bias / Fact Check resource.\footnote{See the Iffy.news page: \href{https://mediabiasfactcheck.com/infowars-alex-jones/}{https://mediabiasfactcheck.com/infowars-alex-jones/}.} On May 2, 2019, Facebook announced they would prohibit users from sharing Infowars content unless they are explicitly condemning the material.\footnote{See \href{https://www.wired.com/story/facebook-bans-alex-jones-extremists/}{https://www.wired.com/story/facebook-bans-alex-jones-extremists/}.}

To verify the measure, we used the CrowdTangle API at the ‘/posts/search’ endpoint\footnote{see the documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Search}{github.com/CrowdTangle/API/wiki/Search}.} to collect the 37,242 Facebook public posts that shared a link containing ‘infowars.com’ and that were published between January 1, 2019 and December 31, 2020.\footnote{We found in the collected data some Facebook posts that did not directly share an Infowars link (but rather a YouTube or Facebook video containing an Infowars link in its description), thus we excluded such posts from our data to keep only the 27,721 posts directly sharing an Infowars link.} The command used can be found \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_infowars_data.sh}{here}.

The number of public posts sharing an Infowars link remained globally stable in 2019 (Figure \ref{infowars1} top panel), thus Facebook’s announced ‘ban’ did not appear to prevent users from sharing an Infowars link. Interestingly a clear drop in engagement was observed on May 2, 2019 (Figure \ref{infowars1} bottom panel). The number of reactions per post decreased by -94\% when we compared the two months before and after May 2, 2019, the shares decreased by -96\%, and the comments by -93\%.

\begin{figure}
	\centering

	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/infowars/fb_infowars_1.png}
	\end{multicols}

	\begin{multicols}{1}
		\includegraphics[scale=0.3]{./img/infowars/fb_infowars_2.png} 
	\end{multicols}

	\caption{Public Facebook posts sharing an Infowars link in 2019 and 2020 gathered from the CrowdTangle API. The red line marks the date of May 2, 2019, when Facebook has announced the ban regarding Infowars. (Top panel) Number of daily posts. (Bottom panel) Engagement metrics (average number of reactions, shares and comments per post). }
	\label{infowars1}
\end{figure}

@@

\begin{figure}
	\centering
		\begin{multicols}{1}	
			\includegraphics[scale=0.3]{./img/globalresearch_2021-06-14.png}
		\end{multicols}

		\begin{multicols}{1}
			\includegraphics[scale=0.3]{./img/globalresearch_14_06_2021_16pm_UTC.png} 
		\end{multicols}
		\caption{Screenshots taken on June 14, 2021. }
		\label{fig3}
\end{figure}

@@

\begin{figure*}
\centering
	\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/globalresearch/sum_globalresearch.ca_6_months.jpg} 
		\includegraphics[scale=0.2]{./img/globalresearch/engagement_sum_rolling_1_globalresearch.ca.jpg}
	\end{multicols}

	\begin{multicols}{1}
		\includegraphics[scale=0.2]{./img/globalresearch/tweet.png} 
		\includegraphics[scale=0.2]{./img/globalresearch/warning.png}
	\end{multicols}
\caption{}
\label{fig4}
\end{figure*}

@@

@@

{\color{pink} What about something about recommendations ? authoritative content  }

\subsection{Flags, Notices and labels}

@@

\section{Appendix}
